
---
title: "MENACE"
author: "Aleksandra"
date: "2026-08-31"
categories: [math, interactive]
description: "The hardest part is to understand fundamentals."
title-block-banner: true
include-in-header: libs.html
---
::: {#fig-menace-boxes}
```{r}
#| echo: false
#| warning: false

source("menace_helpers.R")

create_menace_collection()
```

The entire collection of MENACE matchboxes, grouped by turn. Beads in the boxes are used to determine the computer's moves during the game. MENACE plays with crosses and starts the game. Each matchbox represents one board position, with rotated or reflected copies stored only once. 
:::

::: {#fig-menace-move-one}
```{r}
#| echo: false
#| warning: false

render_move1_figure()
```

Move 1: on an empty board, rotations reduce the nine possible opening squares to three distinct choices: a corner, an edge, or the centre. Therefore, at the beginning of the game, the first matchbox has eight beads in each of three colors, 24 beads in total, with each color representing one of these choices. A single bead is drawn and its color determines MENACE's move.
:::

::: {#fig-menace-move-two}
```{r}
#| echo: false
#| warning: false

render_move2_figure()
```

After each player has moved once, the number of distinct next moves depends on which symmetries remain. The three boards correspond to positions #6, #10, and #12 in Figure @fig-menace-boxes. Position #6 has seven distinct choices; positions #10 and #12 each have four because they preserve mirror symmetry.
:::

::: {#fig-menace-move-two-options}
```{r}
#| echo: false
#| warning: false

create_menace_move2()
```

The twelve numbered boxes show every position MENACE can encounter on its second turn.
:::

::: {#fig-menace-reinforcement}
```{r}
#| echo: false
#| warning: false

create_reinforcement_figure()
```

After a game, MENACE updates the bead color it selected in every matchbox used during that game: three beads are added after a win, one after a draw, and one is removed after a loss. The diagram starts from a fresh count of eight beads; other choices in each box are unchanged, and a losing choice never falls below one bead.
:::

::: {#fig-menace-engine}
```{r}
#| echo: false
#| warning: false

render_menace_plot(
  controls_html    = create_menace_controls(),
  board_html       = create_menace_board(),
  curve_html       = create_menace_curve()
)
```

The playable engine chooses among the current matchbox options in proportion to their bead counts. The 10 x 10 area display shows the same probabilities visually, while the learning curve records how MENACE's results change during training.
:::
