
---
title: "MENACE"
author: "Aleksandra"
date: "2026-08-31"
categories: [math, interactive]
description: "The hardest part is to understand fundamentals."
title-block-banner: true
include-in-header: libs.html
---

::: {#fig-menace-move-one}
```{r}
#| echo: false
#| warning: false

source("menace_helpers.R")

render_move1_figure()
```

On an empty board, rotations and reflections reduce the nine possible opening squares to three distinct choices: a corner, an edge, or the centre.
:::

::: {#fig-menace-move-two}
```{r}
#| echo: false
#| warning: false

render_move2_figure()
```

After each player has moved once, the number of distinct next moves depends on which symmetries remain. The three boards correspond to positions #6, #10, and #12 in Figure @fig-menace-boxes. Position #6 has seven distinct choices; positions #10 and #12 each have four because they preserve mirror symmetry.
:::

::: {#fig-menace-boxes}
```{r}
#| echo: false
#| warning: false

htmltools::tagList(
  create_menace_move2(),
  create_menace_collection()
)
```

Each matchbox represents one board position, with rotated and reflected copies stored only once. The five enlarged boxes show every position MENACE can encounter on its second turn; the collection below groups all matchboxes by turn.
:::

::: {#fig-menace-reinforcement}
```{r}
#| echo: false
#| warning: false

create_reinforcement_figure()
```

After a game, MENACE updates the bead color it selected in every matchbox used during that game: three beads are added after a win, one after a draw, and one is removed after a loss. The diagram starts from a fresh count of eight beads; other choices in each box are unchanged, and a losing choice never falls below one bead.
:::

::: {#fig-menace-engine}
```{r}
#| echo: false
#| warning: false

render_menace_plot(
  controls_html    = create_menace_controls(),
  board_html       = create_menace_board(),
  curve_html       = create_menace_curve()
)
```

The playable engine chooses among the current matchbox options in proportion to their bead counts. The 10 x 10 area display shows the same probabilities visually, while the learning curve records how MENACE's results change during training.
:::
