---
title: "Integrals"
author: "Aleksandra"
date: "2026-02-07"
categories: [math, interactive]
image: ""
description: ""
title-block-banner: true
---

## Interactive Integral Visualization

Use the slider below to change the upper limit of integration. The left plot shows the function $f(t) = t$ with the area under the curve shaded. The dashed outline shows the maximum possible area. The right plot shows the value of the integral $\int_0^t \tau \, d\tau = \frac{t^2}{2}$.

```{r}
#| echo: false
#| warning: false

library(plotly)
library(htmlwidgets)

# Load helper functions
source("plot_helpers.R")

# Read external JavaScript file
js_code <- readLines("plot_integral.js", warn = FALSE)
js_code <- paste(js_code, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html <- create_integral_controls(
  id_prefix = "",
  range_min = 0, range_max = 10,
  default_lower = 0, default_upper = 5,
  update_function = "updateIntegralPlot",
  bg_color = "#f8f9fa"
)
plot_containers <- create_plot_containers(id_prefix = "")

# Create minimal plotly widget just to load Plotly.js and run our code
fig <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html, plot_containers)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initIntegralPlot(el);
      }",
      js_code
    )
  )

fig
```

## Logarithmic Function

Now let's look at $f(t) = \ln(t)$. The integral $\int_1^t \ln(\tau) \, d\tau = t \cdot \ln(t) - t + 1$.

```{r}
#| echo: false
#| warning: false

# Read external JavaScript file for log plot
js_code_log <- readLines("plot_log_integral.js", warn = FALSE)
js_code_log <- paste(js_code_log, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html_log <- create_integral_controls(
  id_prefix = "Log",
  range_min = 1, range_max = 10,
  default_lower = 1, default_upper = 5,
  update_function = "updateLogIntegralPlot",
  bg_color = "#f0fff0"
)
plot_containers_log <- create_plot_containers(id_prefix = "Log")

# Create minimal plotly widget just to load Plotly.js and run our code
fig_log <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html_log, plot_containers_log)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initLogIntegralPlot(el);
      }",
      js_code_log
    )
  )

fig_log
```

## Exponential Function

Now let's look at $f(t) = e^t$. The integral $\int_0^t e^\tau \, d\tau = e^t - 1$.

```{r}
#| echo: false
#| warning: false

# Read external JavaScript file for exp plot
js_code_exp <- readLines("plot_exp_integral.js", warn = FALSE)
js_code_exp <- paste(js_code_exp, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html_exp <- create_integral_controls(
  id_prefix = "Exp",
  range_min = 0, range_max = 5,
  default_lower = 0, default_upper = 2.5,
  update_function = "updateExpIntegralPlot",
  bg_color = "#f5f0ff"
)
plot_containers_exp <- create_plot_containers(id_prefix = "Exp")

# Create minimal plotly widget just to load Plotly.js and run our code
fig_exp <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html_exp, plot_containers_exp)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initExpIntegralPlot(el);
      }",
      js_code_exp
    )
  )

fig_exp
```

## Sigmoid Function

The sigmoid function $\sigma(t) = \frac{1}{1+e^{-t}}$ is widely used in machine learning. Its integral is $\int_a^b \sigma(\tau) \, d\tau = \ln(1+e^b) - \ln(1+e^a)$.

```{r}
#| echo: false
#| warning: false

# Read external JavaScript file for sigmoid plot
js_code_sigmoid <- readLines("plot_sigmoid_integral.js", warn = FALSE)
js_code_sigmoid <- paste(js_code_sigmoid, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html_sigmoid <- create_integral_controls(
  id_prefix = "Sigmoid",
  range_min = -5, range_max = 5,
  default_lower = -2, default_upper = 2,
  update_function = "updateSigmoidIntegralPlot",
  bg_color = "#e0ffff"
)
plot_containers_sigmoid <- create_plot_containers(id_prefix = "Sigmoid")

# Create minimal plotly widget just to load Plotly.js and run our code
fig_sigmoid <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html_sigmoid, plot_containers_sigmoid)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initSigmoidIntegralPlot(el);
      }",
      js_code_sigmoid
    )
  )

fig_sigmoid
```
