---
title: "Integrals"
author: "Aleksandra"
date: "2026-02-07"
categories: [math, interactive]
image: ""
description: ""
title-block-banner: true
---

## Interactive Integral Visualization

Use the slider below to change the upper limit of integration. The left plot shows the function $f(t) = t$ with the area under the curve shaded. The dashed outline shows the maximum possible area. The right plot shows the value of the integral $\int_0^t \tau \, d\tau = \frac{t^2}{2}$.

```{r}
#| echo: false
#| warning: false

library(plotly)
library(htmlwidgets)

# Load helper functions
source("plot_helpers.R")

# Read external JavaScript file
js_code <- readLines("plot_integral.js", warn = FALSE)
js_code <- paste(js_code, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html <- create_integral_controls(
  id_prefix = "",
  lower_min = 0, lower_max = 10, lower_default = 0,
  upper_min = 0, upper_max = 10, upper_default = 5,
  update_function = "updateIntegralPlot",
  bg_color = "#f8f9fa"
)
plot_containers <- create_plot_containers(id_prefix = "")

# Create minimal plotly widget just to load Plotly.js and run our code
fig <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html, plot_containers)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initIntegralPlot(el);
      }",
      js_code
    )
  )

fig
```

## Logarithmic Function

Now let's look at $f(t) = \ln(t)$. The integral $\int_1^t \ln(\tau) \, d\tau = t \cdot \ln(t) - t + 1$.

```{r}
#| echo: false
#| warning: false

# Read external JavaScript file for log plot
js_code_log <- readLines("plot_log_integral.js", warn = FALSE)
js_code_log <- paste(js_code_log, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html_log <- create_integral_controls(
  id_prefix = "Log",
  lower_min = 0.1, lower_max = 10, lower_default = 1,
  upper_min = 0.1, upper_max = 10, upper_default = 5,
  update_function = "updateLogIntegralPlot",
  bg_color = "#f0fff0"
)
plot_containers_log <- create_plot_containers(id_prefix = "Log")

# Create minimal plotly widget just to load Plotly.js and run our code
fig_log <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html_log, plot_containers_log)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initLogIntegralPlot(el);
      }",
      js_code_log
    )
  )

fig_log
```

## Exponential Function

Now let's look at $f(t) = e^t$. The integral $\int_0^t e^\tau \, d\tau = e^t - 1$.

```{r}
#| echo: false
#| warning: false

# Read external JavaScript file for exp plot
js_code_exp <- readLines("plot_exp_integral.js", warn = FALSE)
js_code_exp <- paste(js_code_exp, collapse = "\n")

# Create controls and plot containers using helper functions
controls_html_exp <- create_integral_controls(
  id_prefix = "Exp",
  lower_min = 0, lower_max = 5, lower_default = 0,
  upper_min = 0, upper_max = 5, upper_default = 2.5,
  update_function = "updateExpIntegralPlot",
  bg_color = "#f5f0ff"
)
plot_containers_exp <- create_plot_containers(id_prefix = "Exp")

# Create minimal plotly widget just to load Plotly.js and run our code
fig_exp <- plot_ly() %>%
  layout(
    xaxis = list(visible = FALSE),
    yaxis = list(visible = FALSE),
    plot_bgcolor = "white",
    paper_bgcolor = "white",
    height = 1,
    margin = list(l = 0, r = 0, t = 0, b = 0)
  ) %>%
  htmlwidgets::prependContent(htmltools::tagList(controls_html_exp, plot_containers_exp)) %>%
  htmlwidgets::onRender(
    sprintf(
      "function(el, x) {
        el.style.height = '0px';
        el.style.overflow = 'hidden';
        %s
        initExpIntegralPlot(el);
      }",
      js_code_exp
    )
  )

fig_exp
```
